A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-12 kobakian-cook.Rmd
New tas displays

---
title: A Hexagon Tile Map Algorithm for Displaying Spatial Data
author:
  # see ?rjournal_article for more information
  - name: Stephanie Kobakian
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  stephanie.kobakian@monash.edu 
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  dicook@monash.edu 
abstract: >
  This algorithm creates a tessellated hexagon display to represent a set of the spatial polygons. It allocates these hexagon in a manner that preserves the spatial relationship of the geographic units. The algorithm creates a display that showcases spatial distributions, by emphasising the small geographical regions that are often difficult to locate on geographic maps. Spatial distributions have been presented on alternative representations of geography for many years. In modern times, interactivity and animation have begun to play a larger role, as alternative representations have been popularised by online news sites, and atlas websites with a focus on public consumption. Applications are increasingly widespread, especially in the areas of disease mapping, and election results.
preamble: |
  % Any extra LaTeX you need in the preamble
bibliography: kobakian-cook.bib
output: rticles::rjournal_article
editor_options: 
  chunk_output_type: console
---

```{r defaults, echo=FALSE, message=FALSE, warning=FALSE, comment = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      comment = FALSE,
                      echo = FALSE,
                      eval = TRUE,
                      out.width = "90%",
                      retina = 3,
                      dev = "png",
                      fig.pos = "center")
```

@@

library(RefManageR)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")

@@

In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC]. The creation of approximately equal population areas results in dramatically different square meterage of the administrative geographic units, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI]. This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:thyroid-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of melanoma cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Hobart stand out as more orange and red. To create this choropleth map the SA2 polygons for 2011 from the Australian Bureau of Statistics.

@@

```{r choro, fig.cap = "A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), size = 0.1) + 
  scale_fill_identity() + invthm
```

@@

The solutions to this visualisation problem begin with the geography. Alternative maps shift the focus from land area and shape, to the value of the statistics in a group of areas [@ACCAC]. Different styles of cartograms apply transformations to the geographic areas, to highlight the values of the value of the statistic of interest. These displays result in a distortion of the map space to represent differences in the statistic across the areas [@ACCAC] as the statistic of interest is used to determine the cartogram layout.

@@

There are several established alternative visualisation methods. Tile maps, Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:nsw_grid} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, c) a Dorling cartogram and d) a hexagon tile map. Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics of the Australian population.

@@

The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC]. The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In figure \ref{fig:nsw_grid}a) the choropleth map is presented with the alternative maps to allow for comparisons to be made.

```{r fullhexmap, fig.cap = "A choropleth hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
## Create centroids set
centroids <- sa2 %>% 
  create_centroids(., "sa2_name_2011")

@@

## Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.2,
  ## same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities,
  width = 35,
  verbose = FALSE
)

## Prepare to plot
fort_hex <- fortify_hexagon(data = hex_allocated,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.2) %>% 
            left_join(sa2thyroid_ERP %>% select(sa2_name_2011, SIR, p50))

@@

## Make a plot
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
aus_hexmap
```

```{r alternatives}
# Project data 
sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112)
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

###############################################################################
# Contiguous Cartograms

@@

aus_ggcont <- ggplot(cont) +

  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + guides(fill=FALSE)
#aus_ggcont

#ggsave(filename = "figs/aus_ggcont.png", plot = aus_ggcont,
#       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

```{r contcart}

###############################################################################
# Non - Contiguous Cartograms
if (!file.exists("data/ncont.rda")) {
ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()

@@

aus_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill=FALSE)
#aus_ggncont

@@

```{r dorl}

# Non - Contiguous Dorling Cartograms

dorl <- sa2thyroid_ERP %>%
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  invthm + guides(fill = FALSE)
# aus_ggdorl

#ggsave(filename = "figs/aus_ggdorl.png", plot = aus_ggdorl,
#       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

```{r mapgrid, width = 8, height = 10, eval=FALSE}
aus_grid <- ggdraw(xlim = c(0, 2), ylim = c(0, 4)) +
  draw_plot(aus_ggchoro, x = 0, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggcont,  x = 1, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggncont, x = 0, y = 0, width = 1, height = 2) +
  draw_plot(aus_ggdorl,  x = 1, y = 0, width = 1, height = 2)

aus_grid

ggsave(filename = "figs/aus_grid.png", plot = aus_grid,
  device = "png", bg = "transparent", dpi = 300, width = 8, height = 10)
```

```{r tas_displays, width = 10, height = 4}
library(gridExtra)
grid.arrange(aus_ggcont, tas_ggncont, aus_ggdorl, ncol=3)
```

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:nsw_grid}b) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS]. This shown in the northern SA2s of Tasmania with an average (yellow) estimated SIR of Melanoma for males.

@@

The non-contiguous cartogram in figure \ref{fig:nsw_grid}c) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Hobart, shown in the inset map looks reasonable, but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:nsw_grid}d) shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in figure \ref{fig:nsw_grid} a), b) or c).

<!-- When the population is the statistic of interest, the result is a population cartogram. They fail to preserve a recognisable display due to the difference in size of metropolitan and rural areas [@ACTUC], [@GOINO]. -->

<!-- Contiguous cartograms alter the shape of areas, while preserving boundary relationships of neighbours. Non-contiguous cartograms maintain the geographic shape of each geographic area, but will lose the connection to neighbours as the polygon for each geographic area shrinks or grows. -->

@@

There are three key steps performed to create a tessellated hexagon tile map. First derive the set of centroids from the polygons provided, then create the grid of hexagons locations. These two steps are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. To make tessellated plots, the point locations are converted to hexagon shapes, with the hexagon size used in the grid.

@@

Only two inputs are necessary to begin using the algorithm, the shape file, and the id variable. The id column should uniquely identify each geographic unit in the shape file.

@@

```{hexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, they range from much lower than average (blue) to much higher than average (red)."}
aus_hexmap
```

@@

```{r cents}
centroids <- create_centroids(shp_sf = tas_sa2, sf_id = "SA2_NAME16")
```

```{=tex}
\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/1centroids.png}
\caption{\label{fig:centroids_plot}The geographic shapes of the Statistical Areas of Tasmania at Level 2. The points show the locations of the centroids of the SA2 areas.}
\end{figure}
```
#### Step 2: Create the hexagon grid points

@@

```{r grid}
grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)
```

@@

<!-- ```{r grid2, fig.cap = "Grid points to create a tile map.", out.width = "100%", echo = FALSE} -->

<!-- knitr::include_graphics(path = "figs2grid.png") -->

<!-- ``` -->

```{=tex}
\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/2grid.png}
\caption{\label{fig:grid2}Grid points to create a hexagon tile map.}
\end{figure}
```
##### Rolling windows

@@

The closest latitude row and longitude column are found for each centroid location. Then rows and columns of centroids are divided into 20 groups. The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows. The rolling windows can be seen on the bottom and right of the grid shown in Figure \ref{fig:filter-grid}. This will tailor the available grid points to those most likely to be used. It also helps reduce the amount of time taken, as it decreases the amount of points considered for each centroid allocation.

@@

<!-- ```{r filter-grid, fig.cap = "All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots show the grid points left to choose from after the buffer step. The rolling windows show the collections of rows and columns used to filter the hexagon locations.", echo = FALSE, out.width = "100%"} -->

<!-- knitr::include_graphics(path = "figs3grid.png", dpi = 300) -->

<!-- ``` -->

```{=tex}
\begin{figure}[h]
\centering
\includegraphics[width=10cm]{figs/3grid.png}
\caption{\label{fig:filter-grid}The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step.}
\end{figure}
```
##### Centroid to focal point distance

>>>>

---
title: A Hexagon Tile Map Algorithm for Displaying Spatial Data
author:
  # see ?rjournal_article for more information
  - name: Stephanie Kobakian
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  stephanie.kobakian@monash.edu 
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  dicook@monash.edu 
abstract: >
  The algorithm presented creates a display that uses tessellated hexagons to represent a set of spatial polygons. It allocates these hexagon in a manner that preserves the spatial relationship of the geographic units, in light of their positions to points of interest. The display showcases spatial distributions, by emphasising the small geographical regions that are often difficult to locate on geographic maps. Spatial distributions have been presented on alternative representations of geography for many years. In modern times, interactivity and animation have allowed alternative displays to play a larger role. Alternative representations have been popularised by online news sites, and atlas websites with a focus on public consumption. Applications are increasingly widespread, especially in the areas of disease mapping, and election results.
preamble: |
  % Any extra LaTeX you need in the preamble
bibliography: kobakian-cook.bib
output: rticles::rjournal_article
editor_options: 
  chunk_output_type: console
---

```{r defaults, echo=FALSE, message=FALSE, warning=FALSE, comment = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      comment = FALSE,
                      echo = FALSE,
                      eval = TRUE,
                      out.width = "95%",
                      retina = 3,
                      dev = "png",
                      dpi = 300)
```

@@

library(RefManageR)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown", bib.style = "alphabetic", cite.style = 'alphabetic')

@@

In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC]. The creation of approximately equal population areas results in dramatically different square meterage of the administrative geographic units, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI]. This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:choro} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of thyroid cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of dark and light blue areas across the rural communities in all states. The SA2s around Brisbane stand out as more orange and red.

@@

```{r choro, fig.cap = "A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), size = 0.1) + 
  scale_fill_identity() + invthm
aus_ggchoro
```

@@

The solutions to this visualisation problem begin with the geography. Alternative maps shift the focus from land area and shape, to the value of the statistics in a collection of areas [@ACCAC]. Each style of cartogram applies a different transformation to the geographic areas, to highlight the values of the statistic of interest. These displays result in a distortion of the map space to represent features of the distribution across the areas [@ACCAC] as the statistic of interest is used to determine the cartogram layout.

@@

There are several established alternative visualisation methods. Tile maps, Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:tas_displays} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, and c) a Dorling cartogram.

@@

The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC]. The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics within communities of the Australian population. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In figure \ref{fig:tas_displays} the choropleth map can be seen underneath each alternative map display to allow for comparisons to be made.

```{r fullhexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}

if (!file.exists("data/aus_hexmap.rda")) {

## Create centroids set
centroids <- sa2 %>% 
  create_centroids(., "sa2_name_2011")

@@

## Allocate polygon centroids to hexagon grid points
aus_hexmap <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.2,
  ## same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities,
  width = 35,
  verbose = FALSE
)

save(aus_hexmap, file = "data/aus_hexmap.rda") }

load("data/aus_hexmap.rda")

## Prepare to plot
fort_hex <- fortify_hexagon(data = aus_hexmap,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.2) %>% 
            left_join(sa2thyroid_ERP %>% select(sa2_name_2011, SIR, p50))

@@

## Make a plot
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()

```{r alternatives}
# Projected data 
sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112) %>% filter(state_name_2011 == "Tasmania")
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

# Contiguous Cartograms

@@

tas_ggcont <- ggplot(cont) + 
  geom_sf(fill = "lightgrey", colour = "black", data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + guides(fill=FALSE) +
  theme_map() + ggtitle("a")
#tas_ggcont

@@

```{r contcart}
# Non - Contiguous Cartograms

if (!file.exists("data/ncont.rda")) {
ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()

@@

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "lightgrey", colour = "black",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill=FALSE) +
  theme_map() + ggtitle("b")
#tas_ggncont

@@

```{r dorl}
# Non - Contiguous Dorling Cartograms

dorl <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
tas_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "lightgrey", colour = "black",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  guides(fill = FALSE) +
  theme_map() + ggtitle("c")
#tas_ggdorl
```

```{r tas_displays, width = 10, height = 3, fig.cap = "The three displays show alternative maps of the Statistical Areas at Level 2 in the Australian state of Tasmania. These alternatives show the (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania."}
library(gridExtra)
grid.arrange(tas_ggcont, tas_ggncont, tas_ggdorl, ncol=3)
```

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:tas_displays}b) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS]. This shown in the northern SA2s of Tasmania with an average (yellow) estimated SIR of thyroid for males.

@@

The non-contiguous cartogram in figure \ref{fig:tas_displays}c) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Hobart, shown in the inset map looks reasonable, but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:tas_displays}c) shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in figure \ref{fig:tas_displays} a), or b).

@@

The result of the algorithm is a display that emphasises the densely populated geographic regions surrounding the capital cities in Australia.
Each capital city forms it's own layer of connected hexagons. Figure \ref{fig:hexmap} allows the distributions in the capital cities to be viewed at a deeper scale. Sydney is now notable for the considerable amount of red areas, not just for the surrounding orange areas around the New South Wales capital visible in the choropleth display in Figure \ref{fig:choro}. Brisbane and Perth also draw attention for the many orange areas. Interestingly there is no clear divide between rural and urban SA2 areas as Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid cancer diagnosis.

```{r hexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for males, they range from much lower than average (blue) to much higher than average (red)."}
aus_hexmap
```

There are three key steps performed to create the tessellated hexagon tile map shown in Figure \ref{fig:hexmap}. First, derive the set of centroids from the polygons provided, then create the grid of hexagons locations. These two steps are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. To make tessellated plots, the point locations are converted to hexagon shapes, with the hexagon size used in the grid.

@@

Only two inputs are necessary to begin using the algorithm, the shape file, and the id variable. The id variable should uniquely identify each geographic unit in the shape file.

@@

@@

```{r centroids_plot, fig.cap = "The Australian Cancer Atlas data has determined the colour of each Statistical Area of Australian at Level 2. A choropleth map (a) of Standardised Incidence Rates (SIRs) is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented."}
centroids <- create_centroids(
  shp_sf = sa2 %>% filter(state_name_2011 == "Tasmania"), 
  sf_id = "sa2_name_2011")

hex_size <- .2
buffer_dist <- 2

bbox <- tibble(min = c(min(centroids$longitude), min(centroids$latitude)),
  max = c(max(centroids$longitude), max(centroids$latitude)))

grid <- tibble::as_tibble(expand.grid(hex_long = seq(bbox$min[1] - buffer_dist,
  bbox$max[1] + buffer_dist,
  hex_size),
  hex_lat = seq(bbox$min[2] - buffer_dist,
    bbox$max[2] + buffer_dist,
    hex_size)))

fort_tas <- fort_aus %>% filter(state_name_2011 == "Tasmania")

tas_centroids <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude, colour = SIR), size = 3, shape = 19,
             data = st_drop_geometry(sa2thyroid_ERP) %>% 
               select(sa2_name_2011, SIR) %>%  
               left_join(centroids)) +
  scale_colour_identity() +
  theme_void() + 
  coord_equal()

tas_hexmap <- ggplot() +
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR),
               data = fort_hex %>% filter(sa2_name_2011 %in% fort_tas$sa2_name_2011)) +
  scale_fill_identity() +
  theme_void() + 
  coord_equal()

grid.arrange(tas_centroids, tas_hexmap, nrow=1)
```

#### Step 2: Create the hexagon grid points

@@

```{r grid1}
#grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)

g1 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude), data = centroids, colour = "#374926") + theme_void() + 
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = grid, size = 1.5) +
  coord_equal() + ggtitle("a")
```

@@

```{r grid2, fig.cap = "Complete set of grid points to create a tile map.", height = 4}
# Find every second latitude
shift_lat <- grid %>% dplyr::select(hex_lat) %>%
  dplyr::distinct() %>%
  dplyr::filter(dplyr::row_number() %% 2 == 1) %>% unlist()

# Shift the longitude of every second latitude to the right to make hex structure
grid <- grid %>%
  dplyr::mutate(hex_long = ifelse(hex_lat %in% shift_lat, hex_long,
    hex_long + (hex_size / 2))) %>%
  dplyr::mutate(id=1:NROW(.))  %>%
  dplyr::mutate(assigned=FALSE)

g2 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude), data = centroids, colour = "#374926") + theme_void() + 
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = grid, size = 1.5) +
  coord_equal() + ggtitle("b")

grid.arrange(g1, g2, nrow = 1)
```

##### Rolling windows

@@

The closest latitude row and longitude column are found for each centroid location. Then rows and columns of centroids are divided into 20 groups. The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows. The rolling windows can be seen on the bottom and right of the grid shown in Figure \ref{fig:grid3}. This will tailor the available grid points to those most likely to be used. It also helps reduce the amount of time taken, as it decreases the amount of points considered for each centroid allocation.

@@

```{r grid3, fig.cap = "The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step."}

full_grid <- grid <- grid %>%
  mutate(hex_long_int = dense_rank(hex_long)-1,
    hex_lat_int = dense_rank(hex_lat)-1)

nlong <- length(unique(grid$hex_long))
nlat <- length(unique(grid$hex_lat))

  # Add grid row and column values to centroids data set 
  centroids <- centroids %>%
    mutate(
      # long int for integer value of longitude column
      long_int = round((longitude - min(grid$hex_long)) / 
                         (max(grid$hex_long) - min(grid$hex_long)) * nlong, 0),
      # lat int for integer value of latitude column
      lat_int = round((latitude - min(grid$hex_lat)) / 
                        (max(grid$hex_lat) - min(grid$hex_lat)) * nlat, 0)
    )

  # Amount of lats and longs in each group of rows and columns
  lat_size <- round(nlat / 20, 0)
  long_size <- round(nlong / 20, 0)

  # Make a list of the min and max of the groups
  # Effectively creates manual sliding windows
  nlat_list <- purrr::map2(seq(1:nlat), lat_size + seq(1:nlat), c)
  nlong_list <- purrr::map2(seq(1:nlong), long_size + seq(1:nlong), c)

  # LATITUDE ROWS FILTER

  # Function to return the centroids that fall in the latitude window
  lat_window <- function(bounds, cents = centroids, maximum = nlat) {

    max_int <- min(bounds[2], maximum)

    cents_in <- filter(cents, between(lat_int, bounds[1], max_int))
    return(cents_in)
  }

  # Function to return the centroids that fall in the longitude window
  long_window <- function(bounds, cents = centroids, maximum = nlong) {

    max_int <- bounds[2]

    while (max_int > maximum) {
      max_int <- max_int - 1
    }

    cents_in <- filter(cents, between(long_int, bounds[1], max_int))
    return(cents_in)
  }

  # amount of latitude in sliding window
  lat_windows <- purrr::map(.x = nlat_list, .f = lat_window, centroids, nlat)

  # LONGITUDE COLS FILTER
  long_windows <- purrr::map(.x = nlong_list, .f = long_window, centroids, nlong)

  #########################################################
  ###                ROLLING MIN & MAX                  ###

  # DEFINE FUNCTION
  # find the min and max longitude for each latitude
  range_rows <- purrr::map_dfr(
    .x = lat_windows,
    .f = function(data) {
      data %>%
        dplyr::summarise(
          long_min = ifelse(purrr::is_empty(long_int), NA, min(data$long_int)),
          long_max = ifelse(purrr::is_empty(long_int), NA, max(data$long_int))
        )
    }
  )

  # find the min and max longitude for each latitude
  range_cols <- purrr::map_dfr(.x = long_windows, .f = function(data) {
    data %>%
      dplyr::summarise(
        lat_min = ifelse(purrr::is_empty(lat_int), NA, min(data$lat_int)),
        lat_max = ifelse(purrr::is_empty(lat_int), NA, max(data$lat_int))
      )
  })

  # find the min and max longitude for each latitude
  range_rows <- purrr::map_dfr(.x = lat_windows, .f = function(data) {
      data %>%
        dplyr::summarise(
          long_min = ifelse(purrr::is_empty(long_int), NA, min(data$long_int)),
          long_max = ifelse(purrr::is_empty(long_int), NA, max(data$long_int))
        )
    }
  )

  # find the min and max longitude for each latitude
  range_cols <- purrr::map_dfr(.x = long_windows, .f = function(data) {
    data %>%
      dplyr::summarise(
        lat_min = ifelse(purrr::is_empty(lat_int), NA, min(data$lat_int)),
        lat_max = ifelse(purrr::is_empty(lat_int), NA, max(data$lat_int))
      )
  })

  #########################################################
  ###                ROLLING AVERAGES                   ###

  mean_range <- function(bounds, data) {
    data[bounds[1]:min(bounds[2], NROW(data)), ] %>%
      dplyr::summarise(across(ends_with("min"), 
                         ~mean(.x, na.rm = TRUE), .names = "mean_{col}"),
                       across(ends_with("max"), 
                              ~mean(.x, na.rm = TRUE), .names = "mean_{col}")) %>% 
      # in cases where all values are NA mean is Nan
      # make mean value of NaN be NA
      dplyr::summarise(across(starts_with("mean"), ~ifelse(.x == "NaN", NA, .x)))
  }

  # smooth the minimums
  av_range_rows <- purrr::map_dfr(.x = nlat_list, mean_range, range_rows) %>%
    bind_cols(lat_id = c(seq(1:nlat) + lat_size), .)

  # smooth the minimums
  av_range_cols <- purrr::map_dfr(.x = nlong_list, mean_range, range_cols) %>%
    bind_cols(long_id = c(seq(1:nlong) + round(long_size / 2)), .)

  # APPLY A BUFFER
  # change buffer to amount of hexagons (ints) either side
  hex_buffer <- floor(buffer_dist / hex_size)

grid <- grid %>%
    left_join(., av_range_rows, by = c("hex_lat_int" = "lat_id")) %>%
    left_join(., av_range_cols, by = c("hex_long_int" = "long_id")) %>%
    rowwise() %>%
    mutate(long_buffer = ifelse(between(
      hex_long_int, mean_long_min - hex_buffer,
      mean_long_max + hex_buffer), "in", "out")) %>%
    mutate(lat_buffer = ifelse(between(
      hex_lat_int, mean_lat_min - hex_buffer,
      mean_lat_max + hex_buffer), "in", "out")) %>%
    filter(lat_buffer == "in" | long_buffer == "in")
